Sparse spectral and $p$-finite element methods for partial differential equations on disk-slices and trapeziums

### 2019-04-11 papers/halfdisk/halfdisk.tex
Addressed some of SO's points

The motivation for this work is solving partial differential equations on sub-domains of the sphere. In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \sotodo{Add citation to this}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family. Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction. Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids.

@@

For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
\begin{align}
\begin{cases}
(\alpha,\beta) &:= (0.1) \\
(\gamma,\delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= w_H^{(a,b)}(x) := x^a \: (1-x^2)^b \\
w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{cases}
\end{align}
\bstodoinline{Hope the following makes sense. The generalisation I have, and how I originally wrote/coded up the half-disk case use slightly different conventions regarding the parameters - would it be worth reframing the paper to match this generalisation, or is this fine?}
\sotodoinline{Yes I think its fine} 
Note here the we have made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case. This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.

The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\). The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials. Thus we arrive at  a two parameter family of 2D orthogonal polynomials\sotodo{We should probably use a different letter than $P$ to avoid confusion with Jacobi polynomials}
\ \(\{\Pnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}\label{eq:diskpolys}
 \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight
\begin{align}
W^{(a,b)}(x,y) = w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
\end{align}

@@

$J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded\sotodo{Define Banded-Block-sBanded} structures
\begin{align}
J_{x/y}^{(a,b)} &= \begin{pmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align}
where
\begin{align}
A^x_n &:= \begin{pmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^x_n &:= \begin{pmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
\nonumber \\
A^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
C^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
\end{align}

@@

Combining each system in (\ref{eqn:jacobimatricesdefinition}) we can write the block-tridiagonal system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigPab(x,y)
=
\begin{pmatrix}
	P^{(a,b)}_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
\end{pmatrix}.
\end{align}
where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:\sotodo{Remove $A^z_n$?}
\begin{align}
A_n &:= \begin{pmatrix}
		A^x_n \\
		A^y_n \\
		A^z_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+3)} \quad (n \ne N), \quad
C_n := \begin{pmatrix}
		C^x_n \\
		C^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{pmatrix}
		B^x_n \\
		B^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y) := \begin{pmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
\end{align}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{n+3}$\sotodo{n+1?}. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\begin{align}
\begin{cases}
\bigPab_{-1}(x,y) := 0 \\
\bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots.
\end{cases}
\end{align}

@@

We can then obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
\[
\bigWii(x,y) = \Wii(x,y) \: \bigPii(x,y)
\]
to coefficients in the non-weighted space $\bigPii(x,y)$.

Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}. Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$. The matrix operator for the Laplacian we denote $\laplacewii$ acting on the coefficients vector is then given by
\begin{align}
    \laplacewii := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
\end{align}

Further, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space
$
\bigPoo(x,y)
$
to coefficients in the space
$
\bigP^{(2,2)}(x,y).
$
Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}. We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
\begin{align}
    \laplaceoo := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(1,2)\to(2,2)} \: D_y^{(1,1)} \: T^{(0,1)\to(1,1)} \: D_y^{(0,0)}.
\end{align}

We can use the same technique for constructing a sparse Laplacian operator to construct the Biharmonic operator $\Delta^2$, that ensures we use a basis that satisfies both zero Dirichlet and Neumann conditions. On this note, we can further represent the Laplacian as a map from coefficients in the space 
$
\bigW^{(2,2)}
$
to coefficients in the space
$
\bigPoo.
$
Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$. We denote this matrix operator as $\laplacewtt$, and is given by
\begin{align}
	\laplacewtt := W_x^{(1,1)} \: W_x^{(2,2)} + T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)} \: T_W^{(2,1)\to(1,1)} \: W_y^{(2,2)}.
\end{align}
The sparsity and structure of this matrix can be seen in Figure \ref{fig:sparsity}.

The sparsity of the these operators can be shown using a simple integration by parts argument. \sotodoinline{Add argument}

@@

To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
\begin{align}
x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x), \quad n = 0,1,2,\dots
\end{align}
where we define $\beta_{-1}^{(a,b)} = 0$ and $H_{-1}^{(a,b)}(x) \equiv 0$. We use a variant of the Stieltjes algorithm where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature. This has the benefit that it is easier to incorporate high-precision arithmetic, which we use to overcome ill-conditioning present when $b$ is large, as required for large $n$ in \eqref{eq:diskpolys}. The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm.

\begin{algorithm}
\sotodo{This is just Stieltjes right? We can probably delete it and reference Gautschi.}Inputs: max degree $N$, weight $w(x)$
\begin{flalign*}
	&\text{Define } H_0(x) \equiv \frac{1}{\norm{1}_w} \\
	&\text{Define } v(x) = x  H_0(x) \\
	&\text{Set } \alpha_0 = \ip<v, H_0>_w \\
	&\text{Define } \tilde{v}(x) = v(x) - \alpha_0 H_0(x) \\
	&\text{Set } \beta_0 = \norm{\tilde{v}}_w \\
	&\text{Define } H_1(x) = \frac{\tilde{v}(x)}{\beta_0} \\
	&\text{FOR } n = 1:N \\
		& \quad \quad \text{Define } v(x) = x H_n - \beta_{n-1} H_{n-1} \\
		& \quad \quad \text{Set } \alpha_n = \ip<v, H_n>_w \\
		& \quad \quad \text{Define } \tilde{v}(x) = v(x) - \alpha_n H_n(x) \\
		& \quad \quad \text{Set } \beta_n = \norm{\tilde{v}}_w \\
		& \quad \quad \text{Define } H_{n+1}(x) = \frac{\tilde{v}(x)}{\beta_n} \\
	&\text{END }
\end{flalign*}
Outputs: OP set $(H_0, \dots, H_N)$, coefficients $(\alpha_0, \dots, \alpha_N)$ and $(\beta_0, \dots, \beta_N)$
\end{algorithm}

@@

@@

Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$: \sotodo{Add sentence explaining how symmetry is imprtant here}
\begin{align}
\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j) \big], \\
|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
\end{align}

@@

\subsection{Calculating non-zero entries of the operator matrices}

First, note that we can expand $\pddx \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$. The coefficients of the expansion are then the entries of the relevant operator matrix. By a simple integration by parts argument, we have that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$. These are then calculated by taking inner products, i.e. 
\begin{align*}
c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}
\end{align*}
where \sotodo{Why giant $\prime$?}
\begin{align}
	&\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
	&= \ip< x \: H_{n-k}^{(a, b+k+\half)} \prime (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
	& \quad - k \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
	& \quad + \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
		\: \ip< y \: P_k^{(b,b)} \prime (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
\end{align}
We can then use quadrature to calculate the 1D inner products. This is much cheaper than using the 2D quadrature rule.

Further, we can expand $\pddy \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$. The coefficients of the expansion are then the entries of the relevant operator matrix. By a simple integration by parts argument, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$. This is then calculated by taking inner products, i.e. 
\begin{align*}
c_{n-1,k-1} = \frac{\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
\end{align*}
where \sotodo{Why giant $\prime$?}
\begin{align}
	\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{H_0^{(a,b+k+\half)}}^2_{w_H^{(a, b+k+\half)}} 
		\: \ip< P_k^{(b,b)} \prime, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
\end{align}
where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ \sotodo{Aren't these $= 1$ since we orthonormalize?} $\forall m, n = 0, 1, 2, \dots$.

>>>>

The motivation for this work is solving partial differential equations on sub-domains of the sphere. In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \sotodo{Add citation to this} \bstodo{should we instead take the Xu approach?}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family. Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction. Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids.

@@

For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
\begin{align}
\begin{cases}
(\alpha,\beta) &:= (0.1) \\
(\gamma,\delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= w_H^{(a,b)}(x) := x^a \: (1-x^2)^b \\
w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{cases}
\end{align}
Note here the we have made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case. This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.

The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\). The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials. Thus we arrive at  a two parameter family of 2D orthogonal polynomials\sotodo{We should probably use a different letter than $P$ to avoid confusion with Jacobi polynomials}
\ \(\{\Pnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}\label{eq:diskpolys}
 \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight
\begin{align}
W^{(a,b)}(x,y) = w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
\end{align}

@@

$J_x^{(a,b)}, J_y^{(a,b)}$ have Banded-Block-Banded structures. A block-banded matrix is a block $n$-diagonal matrix where $n$ is reasonably small for the size of the full matrix. A banded-block-banded matrix is one where the blocks themselves are banded matrices. $J_x^{(a,b)}, J_y^{(a,b)}$ are given by:
\begin{align}
J_{x/y}^{(a,b)} &= \begin{pmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align}
where
\begin{align}
A^x_n &:= \begin{pmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^x_n &:= \begin{pmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
\nonumber \\
A^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
C^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
\end{align}

@@

Combining each system in (\ref{eqn:jacobimatricesdefinition}) we can write the block-tridiagonal system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigPab(x,y)
=
\begin{pmatrix}
	P^{(a,b)}_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
\end{pmatrix}.
\end{align}
where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
\begin{align}
A_n &:= \begin{pmatrix}
		A^x_n \\
		A^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)} \quad (n \ne N), \quad
C_n := \begin{pmatrix}
		C^x_n \\
		C^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{pmatrix}
		B^x_n \\
		B^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y) := \begin{pmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
\end{align}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{n+2}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\begin{align}
\begin{cases}
\bigPab_{-1}(x,y) := 0 \\
\bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n(x,y) - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots.
\end{cases}
\end{align}

@@

We can then obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
$$
\bigWii(x,y) = \Wii(x,y) \: \bigPii(x,y)
$$
to coefficients in the non-weighted space $\bigPii(x,y)$. Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$. The matrix operator for the Laplacian we denote $\laplacewii$ acting on the coefficients vector is then given by
\begin{align}
    \laplacewii := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
\end{align}
Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.

We can use the same technique for constructing a sparse Laplacian operator to construct the Biharmonic operator $\Delta^2$, that ensures we use a basis that satisfies both zero Dirichlet and Neumann conditions. On this note, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bigPoo(x,y)$ to coefficients in the space $\bigP^{(2,2)}(x,y)$. We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
\begin{align}
    \laplaceoo := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(1,2)\to(2,2)} \: D_y^{(1,1)} \: T^{(0,1)\to(1,1)} \: D_y^{(0,0)}.
\end{align}
Further represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bigPoo$. Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$. We denote this matrix operator as $\laplacewtt$, and is given by
\begin{align}
	\laplacewtt := W_x^{(1,1)} \: W_x^{(2,2)} + T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)} \: T_W^{(2,1)\to(1,1)} \: W_y^{(2,2)}.
\end{align}
We can then construct a matrix operator for $\Delta^2$ that will take coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bigP^{(2,2)}$. Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$. The matrix operator for the Biharmonic operator is then given by
\begin{align}
	\biharmonic = \laplaceoo \: \laplacewtt.
\end{align}
The sparsity and structure of this biharmonic operator is seen in Figure \ref{fig:sparsity}.

The sparsity of the these operators can be shown using a simple integration by parts argument, see Section \ref{subsection:Computation-operatormatrices}.

@@

To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
\begin{align}
x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x), \quad n = 0,1,2,\dots
\end{align}
where we define $\beta_{-1}^{(a,b)} = 0$ and $H_{-1}^{(a,b)}(x) \equiv 0$. We use a variant of the Stieltjes procedure \cite{gautschi1982generating} where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature. This has the benefit that it is easier to incorporate high-precision arithmetic, which we use to overcome ill-conditioning present when $b$ is large, as required for large $n$ in \eqref{eq:diskpolys}. The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm.

@@

Note here that due to the symmetry in the domain, the integral of an odd function in the second variable (i.e. if $f(x, \dot)$ is odd for fixed $x$) over $\Omega$ is zero. Thus, the quadrature rule is such that we only consider the even part of a function.

@@

Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$: 
\begin{align}
	\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j) \big], \\
	|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
\end{align}

@@

\subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}

Since $\{P^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \Pnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$. The coefficients of the expansion are then the entries of the relevant operator matrix. We can use an integration by parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$. First, note that
\begin{align*}
c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
\end{align*}
Second, note that
\begin{align}
	\rho'(x) = -x \: \rho(x)^{-2}. \label{eqn:rhoderivative}
\end{align}
Now, using a change of variables and (\ref{eqn:rhoderivative}), we have that
\begin{align}
	&\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
	&= \ip< x \: H_{n-k}^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
	& \quad - k \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
	& \quad + \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
\end{align}
We can see that, from each term's second factor as written, the above is zero for $j < k-2$. Now, using integration by parts and noting that the weight $w_H^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
\begin{align*}
	&\ip< x \: H_{n-k}^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}}  \\
	&= \ip<H_{n-k}^{(a, b+k+\half) \: \prime}, \: w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half)} >_1 \\
	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \ddx \Big[w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half)} \Big] >_1 \\
	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \Big[(a+1) \: w_H^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: w_H^{(a+2, b+\half(k+j+1))} \Big] \: H_{m-j}^{(a, b+1+j+\half)} >_1 \\
		&\quad \quad \quad + \ip<H_{n-k}^{(a, b+k+\half)}, \: w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half) \: \prime}>_1 \\
	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \Big[(a+1) \: w_H^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: w_H^{(2, \half(j-k))} \Big] \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+k+\half)}} \\
		&\quad \quad \quad + \ip<H_{n-k}^{(a, b+k+\half)}, \: w_H^{(1, 1+\half(j-k))} \: H_{m-j}^{(a, b+1+j+\half) \: \prime}>_{w_H^{(a, b+k+\half)}}
\end{align*}
which is zero for $m < n - 2$, and further that
\begin{align*}
